---
title: "Investigating positron emission tomography (PET) scan use in diffuse large B-cell lymphoma (DLBCL) - A Hospital Episode Statistics feasibility analysis"
author: "Andre Ng"
date: "`r Sys.Date()`"
output: html_document
format: html
editor: source
subtitle: "Cohort Creation"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


# renv::snapshot()
#Check for any packaging issues
renv::status()
# renv::restore()


```



```{r load_libraries}
library(coRanalysis)
library(dplyr)
library(dbplyr)
library(tidyr)
library(readxl)

source(here::here("R", "setup.R"))


#Load HES inpatient and APC data (outpatient)
hes_con <- connect_to_db("hesdata_processed_hiq")

inpatient <- tbl(hes_con, in_schema("public", "inpatient"))
outpatient <- tbl(hes_con, in_schema("public", "outpatient"))


#Combine diagnoses and procedures from inpatient and outpatient data
hes_events <- pivot_hes(inpatient, outpatient)


#Transfer codelist from protocal and upload to database
kite_codelist <- read_excel(here::here("kite_codelist.xlsx"))

cprd_con <- connect_to_db(database = "cprd")
copy_to(cprd_con, kite_codelist, "kite_codelist",  overwrite = FALSE)

# db$upload(kite_codelist, "kite_codelist")

#Load in codelist for kite


```


# Attrition Table Overview
-Patients identified with a code for DLBCL between 1 January 2021 and 30 April 2023
-Patients with DLBCL in the eligibility period with complete information (gender and DOB)
-Patients with DLBCL in the eligibility period with complete information aged ≥18 years at diagnosis
-Patients with DLBCL in the eligibility period with complete information, aged ≥18, without evidence of HSCT prior to 31 December 2020
-"STUDY POPULATION: Patients with DLBCL in the eligibility period and complete information, aged ≥18, without evidence of HSCT or any chemotherapy prior to 31 December 2020"


# Inclusion
1.	A diagnosis of DLBCL (ICD-10 coding C833 Diffuse large B-cell lymphoma) in the APC or OP datasets


```{r}
# dlbcl_cohort <- 
#   kite_codelist %>%
#     filter(Variable == "Diagnosis") %>%
#     rename(code = Code,
#            code_type = `Code type`,
#            term = Description,
#            category = Variable) %>%
#     inner_join(hes_events, join_by(code ,code_type)) %>%
#     select(-category) %>%
  
dlbcl_cohort <- 
  hes_events %>%
    filter(code == "C833",
           code_type == "icd10") %>%
    ###
    filter(between(event_date, lookback_start, study_end)) %>% #within data period
    # Get first event for each patient
    group_by(token_person_id) %>%
    slice_min(event_date, n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    mutate(
      token_person_id,
      index_dlbcl_date = event_date,
      index_dlbcl_code = code,
      index_dlbcl_key = key,
      index_dlbcl_key_type = key_type,
      .keep = "none"
    ) %>%
    compute()

#Total patients with DLBLCL from lookback date start to end of study period (within data period)
dlbcl_cohort %>%
  count() #40765

#Patients identified with a code for DLBCL between 1 January 2021 and 30 April 2023
dlbcl_cohort %>%
  filter(between(index_dlbcl_date, lookback_start, eligible_end)) %>%
  count() #37370


```


# Exclusion
1.	≥18 years at diagnosis
2.	Evidence of DLBCL or haematopoietic stem cell transplantation (HSCT) prior to 31 December 2020 (see codelist in Appendix)
3.	Evidence of treatment with any chemotherapy within HES prior to 31 December 2020
4.	Patients with missing gender or date of birth. 


6 months follow-up?
complete data linkage availability (including demographics?)

```{r}
demographics <- list()

demographics$ip <- inpatient %>%
  select(token_person_id, index_dlbcl_key = epikey, index_dlbcl_age = startage, index_dlbcl_sex = sex) %>%
  mutate(index_dlbcl_type = "epikey", index_dlbcl_key = as.character(index_dlbcl_key))

demographics$op <- outpatient %>%
  select(token_person_id, index_dlbcl_key = attendkey, index_dlbcl_age = apptage, index_dlbcl_sex = sex) %>%
  mutate(index_dlbcl_key_type = "attendkey", across(c(index_dlbcl_age, index_dlbcl_sex), as.integer))

index_dlbcl_demographics <- purrr::reduce(demographics, dplyr::union_all)


```


