---
title: "Investigating positron emission tomography (PET) scan use in diffuse large B-cell lymphoma (DLBCL) - A Hospital Episode Statistics feasibility analysis"
author: "Andre Ng"
date: "`r Sys.Date()`"
output: html_document
format: html
editor: source
subtitle: "Baseline Characteristics"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


# renv::snapshot()
#Check for any packaging issues
renv::status()
# renv::restore()


```

```{r load_libraries_and_datasets}
library(coRanalysis)
library(dplyr)
library(dbplyr)
library(tidyr)
library(readxl)

source(here::here("R", "setup.R"))
source(here::here("R", "util.R"))

#Load in codelist for kite
kite_codelist <- tbl(db_con, in_schema("kite", "kite_codelist"))


#Load in cohort dataset
dlbcl_cohort <- tbl(db_con, in_schema("kite", "dlbcl_cohort"))
  
#Load HES inpatient and APC data (outpatient)
inpatient <- tbl(db_con, in_schema("public", "inpatient"))
outpatient <- tbl(db_con, in_schema("public", "outpatient"))

#Combine diagnoses and procedures from inpatient and outpatient data
hes_events <- pivot_hes(inpatient, outpatient)

#Load in deaths table
deaths <- tbl(db_con, in_schema("kite", "dlbcl_cohort_deaths"))

```

```{r setup}
# Individual baseline characteristics - One row per pt
pt_char <- list()

# Cohort characteristics - Aggregated and formatted for output
bs_char <- list()


cohort <- dlbcl_cohort %>%
  filter(eligible_flag == 1) %>%
  select(token_person_id,
         index_dlbcl_date,
         index_dlbcl_key,
         index_dlbcl_key_type,
         index_dlbcl_age,
         index_dlbcl_sex,
         index_dlbcl_admidate,
         index_dlbcl_imd,
         source) %>%
  compute()

```


# N (%) of patients with chemotherapy

```{r n_pct_chemotherapy}
#Number of dlbcl patients with a diagnosis code for chemotherapy identified during the eligibility period.

cohort <- 
  hes_events %>%
    # mutate(eligible_chemotherapy_flag = 1) %>% ---------------------------------------------------------------THIS IS IF EXCLUDE CHEMO BEFORE ELIGIBLE PERIOD
    filter(between(event_date, eligible_start, eligible_end)) %>%
    right_join(kite_codelist %>%
                 filter(category == "Chemotherapy"),
               join_by(code_type, code),
               copy = TRUE) %>%
    right_join(cohort,
               join_by(token_person_id),
               copy = TRUE) %>%
    group_by(token_person_id) %>%
    slice_min(event_date, n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    # mutate(eligible_chemotherapy_flag = if_else(is.na(eligible_chemotherapy_flag) == TRUE, 0, 1)) %>% --------THIS IS IF EXCLUDE CHEMO BEFORE ELIGIBLE PERIOD
    rename(
      chemo_date = event_date,
      chemo_code_type = code_type,
      chemo_code = code
    ) %>%
    select(-key,
           -code_index,
           -key_type,
           -term,
           -category) %>%
    compute()

# ---------------------------------------------------------------------------------------------------------THIS IS IF EXCLUDE CHEMO BEFORE DLBCL DIAGNOSIS DATE
cohort <- 
  cohort %>%
    mutate(eligible_chemotherapy_flag = if_else(
      (is.na(chemo_date) == FALSE) &
      (chemo_date >= index_dlbcl_date),
      1,
      0))


pt_char$n_pct_chemotherapy <- cohort

bs_char$n_pct_chemotherapy <- 
  pt_char$n_pct_chemotherapy %>%
    count(eligible_chemotherapy_flag) %>% #9348 or 8012
    collect() %>%
    mutate(n = format_n_pct(n, sum(n))) %>%
    rename(stat = n)


```

# Time between diagnosis and first chemo (internal adhoc request)

```{r}

cohort %>%
  mutate(diag_chemo_years = (datediff("day", index_dlbcl_date, chemo_date)) / 365.25) %>%
  standard_summary(diag_chemo_years)


```


# Create index date
-Patients will be indexed on the date of their first recorded chemotherapy administration after DLBCL diagnosis or, if no chemotherapy is recorded, date of diagnosis of DLBCL.

-Where patients are diagnosed [with dlbcl?] during inpatient admissions, the admission date (‘admidate’) for the admission [with dlbcl] will be used as the date of DLBCL diagnosis to ensure any chemotherapy administered during the same admission is captured. 

-Patients with chemotherapy dates BEFORE DLBCL diagnosis, use chemotherapy date instead?? --------------------ASK JEN

-If the first recorded chemotherapy is ≥6 weeks after the diagnosis, the patient will be indexed on their diagnosis date [of dlbcl]. 


```{r create_index_date}
#----------------------------------------------------------------------------------------------------------THIS IS IF EXCLUDE CHEMO BEFORE ELIGIBLE PERIOD
cohort <-
  cohort %>%
    arrange(token_person_id) %>%
      #Inpatient patients use admidate instead of index_dlbcl_date
      mutate(index_date = if_else(
        (source == "inpatient"), 
            index_dlbcl_admidate, 
            index_dlbcl_date)) %>%
      #Chemo patients with chemodate <6 weeks after dlbcl_date use chemodate, else use index_dlbcl_date or index_dlbcl_admidate
      mutate(index_date = if_else(
        ((eligible_chemotherapy_flag == 1) &
        (datediff("day", index_dlbcl_date, chemo_date) < 42)),
            chemo_date,
            index_date))


```


# Create follow-up end date for each patient

Follow up will end at the earliest of inpatient death or the end of the study period 30 November 2023 (or the latest available data). Inpatient death here refers to people whose death we observe during a hospital admission recorded in APC. 

```{r create_end_date}

cohort <-
  cohort %>%
    left_join(deaths, join_by(token_person_id)) %>%
    mutate(end_date = if_else(
             (is.na(death_date) == FALSE),
             death_date,
             as.Date("2023-11-30")),
           death_flag = if_else(
             (is.na(death_date) == FALSE),
             1,
             0)
          )

#SAVE TABLE?----------------------------------------------------------------------------  


```

# Follow-up time

```{r follow_up_time}

pt_char$followup <-
  cohort %>%
    mutate(follow_up_years = (datediff("day", index_date, end_date)) / 365.25)

bs_char$followup <-
  cohort %>%
  mutate(follow_up_years = (datediff("day", index_date, end_date)) / 365.25) %>%
  standard_summary(follow_up_years)

```

# Age
```{r age}

bs_char$age <-
  cohort %>%
  standard_summary(index_dlbcl_age)

```

# Gender
```{r}
bs_char$gender <-
  cohort %>%
  select(index_dlbcl_sex) %>%
  count(index_dlbcl_sex) %>%
  collect() %>%
  mutate(index_dlbcl_sex = factor(index_dlbcl_sex, c(1,2,0,9), c("Male", "Female", "Unknown", "Unknown"))) %>%
  fill_n(index_dlbcl_sex) %>%
  mutate(n = format_n_pct(n, sum(n)))
  # pivot_stat(index_dlbcl_sex) %>%
  # select(Male, Female, Unknown)

```

# IMD
```{r}

bs_char$deprivation <- 
  cohort %>%
    collect() %>%
    mutate(index_dlbcl_imd = (as.numeric(index_dlbcl_imd) + 1) %/% 2) %>% #use remainder division to convert deciles to quintiles
    summarise_categorical("index_dlbcl_imd") %>%
    rename(stat = Count_Percent) %>%
    arrange(as.numeric(stat))
    # pivot_wider(names_from = index_dlbcl_imd, values_from = stat)


```


