---
title: "Investigating positron emission tomography (PET) scan use in diffuse large B-cell lymphoma (DLBCL) - A Hospital Episode Statistics feasibility analysis"
author: "Andre Ng"
date: "`r Sys.Date()`"
output: html_document
format: html
editor: source
subtitle: "Baseline Characteristics"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


# renv::snapshot()
#Check for any packaging issues
renv::status()
# renv::restore()


```

```{r load_libraries_and_datasets}
library(coRanalysis)
library(dplyr)
library(dbplyr)
library(tidyr)
library(readxl)

source(here::here("R", "setup.R"))
source(here::here("R", "util.R"))

#Load in codelist for kite
kite_codelist <- tbl(db_con, in_schema("kite", "kite_codelist"))


#Load in cohort dataset

#Load in hes_events table

#--------------------------------------------------------------------------------------CURRENTLY THIS SCRIPT IS DEPENDENT ON 01_cohort_creation (ie run that first) because tables cannot be uploaded to database.

```

```{r setup}
# Individual baseline characteristics - One row per pt
pt_char <- list()

# Cohort characteristics - Aggregated and formatted for output
bs_char <- list()


cohort <- dlbcl_cohort %>%
  filter(eligible_flag == 1) %>%
  select(token_person_id,
         index_dlbcl_date,
         index_dlbcl_key,
         index_dlbcl_key_type,
         index_dlbcl_age,
         index_dlbcl_sex,
         index_dlbcl_admidate,
         source) %>%
  compute()

```


# N (%) of patients with chemotherapy

```{r n_pct_chemotherapy}
#Number of patients with a diagnosis code for DLBCL identified during the eligibility period.

pt_char$n_pct_chemotherapy <- 
  hes_events %>%
    mutate(eligible_chemotherapy_flag = 1) %>%
    filter(between(event_date, eligible_start, eligible_end)) %>%
    right_join(kite_codelist %>%
                 filter(category == "Chemotherapy"),
               join_by(code_type, code),
               copy = TRUE) %>%
    right_join(cohort,
               join_by(token_person_id),
               copy = TRUE) %>%
    group_by(token_person_id) %>%
    slice_min(event_date, n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    mutate(eligible_chemotherapy_flag = if_else(is.na(eligible_chemotherapy_flag) == TRUE, 0, 1)) %>%
    rename(
      chemo_date = event_date,
      chemo_code_type = code_type,
      chemo_code = code
    ) %>%
    select(-key,
           -code_index,
           -key_type,
           -term) %>%
    compute()


bs_char$n_pct_chemotherapy <- 
  pt_char$n_pct_chemotherapy %>%
    count(eligible_chemotherapy_flag) %>% #9348
    collect() %>%
    mutate(n = format_n_pct(n, sum(n))) %>%
    rename(stat = n)





```


































# Create index date
Patients will be indexed on the date of their first recorded chemotherapy administration after DLBCL diagnosis or, if no chemotherapy is recorded, date of diagnosis of DLBCL. 
If the first recorded chemotherapy is ≥6 weeks after the diagnosis, the patient will be indexed on their diagnosis date [of dlbcl?]. 
Where patients are diagnosed [with dlbcl?] during inpatient admissions, the admission date (‘admidate’) for the admission [with dlbcl?] will be used as the date of DLBCL diagnosis to ensure any chemotherapy administered during the same admission is captured. 
